---
output: github_document
---

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/linelist)](https://cran.r-project.org/package=linelist)

# Welcome to the *linelist* package!

This is just a template. Please remove, add or edit parts as you see fit.

## Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval = FALSE}
install.packages("linelist")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:
```{r install2, eval = FALSE}
devtools::install_github("reconhub/linelist")
```

Note that this requires the package *devtools* installed.


# What does it do?

The main features of the package include:



# Resources

## Vignettes

An overview and examples of *linelist* are provided in the vignettes:

...

## Websites

The following websites are available:

...

## Getting help online

Bug reports and feature requests should be posted on *github* using the [*issue*](http://github.com/reconhub/linelist/issues) system. All other questions should be posted on the **RECON forum**: <br>
[http://www.repidemicsconsortium.org/forum/](http://www.repidemicsconsortium.org/forum/)

Contributions are welcome via **pull requests**.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

