---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Basic usage
Let us consider some messy `data.frame` as a toy example: 

```{r toy_data}
library(linelist)
ll_sample <- messy_data()
```

We can use `clean_data()` with defaults to let the package to do its best to clean up the data.

```{r}
clean_data(ll_sample)
```

## Intermediate use
The `clean_data()` function comes with switches for many of the cleaning capabilities so that you can define a bespoke cleaning process.

```{r}
clean_data(ll_sample, guess_dates = TRUE, error_tolerance = 0.1)
```

We can prevent the cleaning process from operating on certain columns.

```{r}
clean_data(ll_sample, guess_dates = TRUE, protect = names(ll_sample) == "messy/dates")
```

If we have values that are multi-lingual or possibly have typos, we can generate a wordlist that contains two columns, `from` and `to`, that will be used to transform any found values in specified columns.

```{r}
wordlist <- data.frame(
 from  = c("hopsital", "hopital",  "medical", "feild"),
 to    = c("hospital", "hospital", "clinic",  "field"),
 variables = rep("location", 4),
 stringsAsFactors = FALSE
)

# location data with mis-spellings, French, and English.
messy_locations <- c("hopsital", "h\u00f4pital", "hospital", 
                     "m\u00e9dical", "clinic", 
                     "feild", "field")
llsample_spell <- messy_data()
llsample_spell$location <- factor(sample(messy_locations, 20, replace = TRUE))

clean_data(llsample_spell, wordlists = wordlist, spelling_vars = "variables" )
```

## Advanced usage
As well as the configurable `clean_data()` function, many of the capabilities are exposed individually, supporting data pipelines.

```{r}
llsample_pipe <- ll_sample
llsample_pipe <- clean_dates(llsample_pipe)
llsample_pipe <- clean_variable_names(llsample_pipe)
```

This gives you highly granular control over the process and is advised when building functions for processing data in your specific usecase / environment.

You can then see the impact of changes by comparing two linelist data.frames.

```{r}
compare_data(ll_sample, llsample_pipe)
```