---
title: "linelist: package overview"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*linelist* provides a framework for sensible data standardisation for tabular
linelist data commonly used in outbreak analysis. This includes functionality
for guessing dates, cleaning non-standard variable names, and defining a data
dictionary.

# Installing the package

To install the current stable, CRAN version of the package, type:

```{r install, eval=FALSE}
install.packages("linelist")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:

```{r install2, eval=FALSE}
devtools::install_github("reconhub/linelist")
```

Note that this requires the package *devtools* installed.

# Main visible functions of the package

The main functions of the package include:

 - `clean_data()` will standardise variable names by removing spaces and extra
   punctuation, attempt to parse dates with a given tolerance, and convert all
   characters to the closest ASCII approximation
 - `as_linelist()` will convert a data frame to a linelist object, allowing the
   user to define a data dictionary.
 - `set_dictionary()` defines valid variables to be used in the data dictionary
 - `get_dictionary()` accesses valid variables to be used in the data dictionary
 - `get_epivars()` accesses a variable from the linelist
 - `set_epivars()` adds new dictionary definitions to the linelist
 
The linelist data class allows you to define variables that are important to
your analysis that make it easy to create situation report templates no matter
what the data class is.

As an example, we can use data from the [outbreaks 
package](https://cran.r-project.org/package=outbreaks). 

```{r load_data}
library('outbreaks')
library('incidence')
library('linelist')
# define the important variables
## set_dictionary(
##   "id",
##   "date_onset",
##   "gender",
##   "age",
##   "outcome", 
##   "date_outcome", 
##   "date_hospitalisation" 
##   )

# generate a template linelist with `template_linelist()`
template_linelist(outbreaks::fluH7N9_china_2013)
# convert the data set to a linelist and define the variables
ll <- as_linelist(outbreaks::fluH7N9_china_2013,
  outcome      = "outcome", 
  date_outcome = "date_of_outcome", 
  id           = "case_id",
  date_onset   = "date_of_onset",
  gender       = "gender",
  age          = "age",
  NULL  # don't delete me
)
```

The resulting linelist object acts the same as a data frame in every way, but
it contains a data dictionary in the metadata:

```{r compare_df}
# These are the same
head(ll)
head(outbreaks::fluH7N9_china_2013)

list_epivars(ll) # show available epivars
list_epivars(ll, simple = TRUE) # only names of available epivars
list_epivars(ll, full_dict = TRUE) # ignore non-epivars columns


```

You can use this data dictionary to access variables:

```{r get_stuff}
head(gender(ll))
head(get_epivars(ll, "outcome")) # only "outcome"
```

This becomes useful in generating templates where the columns from the incoming
data may not be standardised. For example, if you want to generate an incidence
curve for the data stratified by gender, you could write something like this:

```{r, ic_gender}
ic <- incidence(date_onset(ll), interval = "week", groups = gender(ll))
plot(ic)
```

Now, if you had a new data set coming into the template, you are prepared:

```{r, new_data}
old_ll <- ll
ll <- as_linelist(outbreaks::mers_korea_2015$linelist,
  "id"         = "id",
  "age"        = "age",
  "gender"     = "sex",
  "date_onset" = "dt_onset"
  )
ic <- incidence(date_onset(ll), interval = "week", groups = gender(ll))
plot(ic)
```

