---
title: "linelist: package overview"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*linelist* provides a framework for sensible data standardisation for tabular
linelist data commonly used in outbreak analysis. This includes functionality
for guessing dates, cleaning non-standard variable names, and defining a data
dictionary.

# Installing the package

To install the current stable, CRAN version of the package, type:

```{r install, eval=FALSE}
install.packages("linelist")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:

```{r install2, eval=FALSE}
devtools::install_github("reconhub/linelist")
```

Note that this requires the package *devtools* installed.

# Main visible functions of the package

The main functions of the package include:

 - `clean_data()` will standardise variable names by removing spaces and extra
   punctuation, attempt to parse dates with a given tolerance, and convert all
   characters to the closest ASCII approximation
 
TODO: FILL THIS IN LATER
